The previous chapter introduced a method for adapting dynamics when some regions of the source and target dynamics are similar and others are not. This method allowed us to adapt a dynamics model and learn reliability in the real world. This removed the need for specialized methods for collecting real-world training data from specific regions of dynamics.

However, for all the dual-arm DOO manipulation experiments thus far, we have assumed the robot is already grasping the object, and that these grasps cannot change during the manipulation. This assumption severely limits the tasks we can do, and in some cases makes the way to do the task inefficient compare to an approach that allows regrasping. This chapter begins to relax this assumption and use regrasping. Although I am not the first to consider regrasping deformable objects (see Section \ref{Proposed:sec:related_work} for related work), there are very few works that focus on regrasping DOOs, and they are not widely applicable to different robot morphologies or tasks. Using the proposed method, I enable robots to re-grasp when stuck, enabling new tasks and increasing robustness.

% By considering regrasping in the manipulation planning problem, we are potentially introducing several new discrete variables into the action space, which could cause exponential growth in the search space and make long horizon planning intractable. Even when tractable, we still face the problem that long horizon planning with unreliable dynamics models gives unsuccessful plans. To avoid this, we leverage tools for topological path planning to prune the search space and avoid expensive and potentially inaccurate rollouts of the dynamics. We propose a novel topological signature for 

\input{Chap5/sections/01_introduction.tex}
\input{Chap5/sections/02_defining_signature}
\input{Chap5/sections/03_illustrative_examples}
\input{Chap5/sections/04_manipulation_with_signature}
\input{Chap5/sections/05_applications}
\input{Chap5/sections/06_conclusion}
